CompDec - Detection of the Compression Algorithm of File Fragments using Convolutional Neural Networks

### 2020-12-31 paper/main.tex
Resolve issues post review

\begin{abstract}
    Conventional techniques in the domain of digital forensics and data carving focus on simple heuristics. Recently, machine learning has been applied to the subject, achieving state-of-the-art results. This research has, however, focused on general file support. No research has been done on the identification of the algorithm used to compress file fragments. A dataset with a sole focus on compression algorithms was developed and released, using GovDocs1 as a base. Using NIST's Statistical Test Suite, it was found that several compression algorithm produce output that are seemingly random - resulting in a difficult classification problem. By training a convolutional neural network on the dataset, an accuracy of 41\% was achieved, highlighting the difficult nature of this problem. The files created by the tools compress, lz4 and bzip2 were accurately classified whilst others such as zip, rar and gzip produced random guesses. Future work could focus on developing a new purpose-built convolutional neural network, or exploring long short-term memory networks.
\end{abstract}

@@

In digital forensics, \textit{data carving} is the act of extracting files directly from some memory media - without any metadata or known filesystem. Conventional techniques use simple heuristics such as magic numbers, headers etc. These techniques do not scale well due to a limited number of supported file types, slow processing speeds and insufficient accuracy.

Recently, machine learning has been applied to the subject, achieving state-of-the-art results both in terms of scale, accuracy and speed. These techniques utilize an efficient feature extraction from files that can be turned into a small image or other representation of the features. The images are then fed to convolutional neural networks to learn to identify parts of files.

These techniques focus on generality to identify files such as documents (.txt, .docx, .ppt, .pdf) and images (.jpg, .png). There is a gap in research when it comes to accurately identifying compressed files and what algorithm was used to compress them. Compression algorithms seek to make data as dense as possible, which will in turn likely yield a higher entropy than a typical file. This in theory could make detection much harder\cite{PENROSE2013}.

@@

In order to answer how compressed files compare to random data in terms of entropy, only some files and random samples may be feasibly studied. To limit the scope, the analyzed files were selected based on some conditions. These conditions are discussed in greater detail in \ref{method:dataset} and \ref{method:entropy}.

@@

\subsection{Compression algorithm selection}

@@

\subsection{Analysis of entropy}
\label{method:entropy}

@@

\section{Conclusion}

@@

\newpage
\bibliographystyle{IEEEtran}
\bibliography{bibliography.bib}

>>>>

\begin{abstract}
    Conventional techniques in the domain of digital forensics and data carving focus on simple heuristics. Recently, machine learning has been applied to the subject, achieving state-of-the-art results. The research, however, has focused on general file support. No research has been done on the identification of the algorithm used to compress file fragments. A dataset with a sole focus on compression algorithms was developed and released, using GovDocs1 as a base. Using NIST's Statistical Test Suite, it was found that several compression algorithm produce output that are seemingly random - resulting in a difficult classification problem. By training a convolutional neural network on the dataset, an accuracy of 41\% was achieved, highlighting the difficult nature of this problem. The files created by the tools compress, lz4 and bzip2 were accurately classified whilst others such as zip, rar and gzip produced random guesses. Future work could focus on developing a new purpose-built convolutional neural network, or exploring long short-term memory networks.
\end{abstract}

@@

In digital forensics, \textit{data carving} is the act of extracting files directly from some memory media - without any metadata or known filesystem. Conventional techniques use simple heuristics such as magic numbers, headers etc. These techniques do not scale well due to a limited number of supported file types, slow processing speeds as well as an insufficient accuracy\cite{chen2018}.

Recently, machine learning has been applied to the subject, achieving state-of-the-art results both in terms of scale, accuracy and speed. These techniques utilize an efficient feature extraction from files that can be turned into a small image or other representation of the features. The images are then fed to Convolutional Neural Networks (CNNs) to learn to identify parts of files.

These techniques focus on generality to identify files such as documents (.txt, .docx, .ppt, .pdf) and images (.jpg, .png). There is a gap in research when it comes to accurately identifying compressed files and what algorithm was used to compress them. Compression algorithms seek to make data as dense as possible, which will in turn likely yield a higher entropy than a typical file. This in theory could make the identification of the algorithms more difficult than typical image classification problems\cite{PENROSE2013}.

@@

To limit the scope of the comparison of compressed files and random data, a subset of all available compression algorithms and types of random data will be considered. The analyzed files were selected based on some conditions. These conditions are discussed in greater detail in \ref{method:dataset} and \ref{method:entropy}.

@@

\subsection{Compression Algorithm Selection}

@@

\subsection{Analysis of Entropy}
\label{method:entropy}

@@

\newpage
\section{Conclusion}

@@

\bibliographystyle{IEEEtran}
\bibliography{bibliography.bib}
